Work Fragmentation in Developer Interaction Data

### 2016-07-07 paper_2.tex
Review 1.6
Reviews 1.7, 1.8

%\RR{For each research question, add a sentence explaining why we look into it}

@@

   & none & $\leq 3$ & $[4 - 5]$ & $[6 - 10]$ & $\geq 11$  \\  
  \hline
  \multicolumn{6}{c}{\textbf{Edits}} \\
  \hline
  mean & 6.29 &	2.59 & 1.55 & 1.29 & 0.91  \\ 
  % \cline{3-6} 
  %t.test & $\hookrightarrow$& \multicolumn{4}{c}{$<$ 2.2e-16} \\
   \cline{3-6} 
  U-test (adjusted) & $\hookrightarrow$ & \multicolumn{4}{c}{$<$ 2.2e-16} \\
  \cline{3-6} 
  Cliff's delta & $\hookrightarrow$	& \textbf{0.61} & \textbf{0.77} & \textbf{0.80} & \textbf{0.87}    \\
  \hline

  \multicolumn{6}{c}{\textbf{Selections}} \\
  \hline 
  mean & 4.73 &	2.93 & 2.26 & 1.79 & 1.54  \\ 
    % \cline{3-6} 
    %t.test & $\hookrightarrow$& 6.258e-08 & 1.623e-12 & 4.87e-16 & $<$ 2.2e-16 \\
     \cline{3-6} 
    U-test (adjusted) & $\hookrightarrow$ & 8.1e-13 & \multicolumn{3}{c}{$<$ 4.4e-16} \\

@@

  \multicolumn{6}{c}{\textbf{Edit ratio}} \\
  \hline 
  mean & 0.55 & 0.39 & 0.33 & 0.34 & 0.33 \\ 
  \cline{3-6} 
%    t.test & $\hookrightarrow$& 1.62E-13 & \multicolumn{3}{c}{$<$ 2.2e-16} \\
     \cline{3-6} 
    U-test (adjusted) & $\hookrightarrow$ & 4.2e-14& \multicolumn{3}{c}{$<$ 2.9e-16} \\
    \cline{3-6} 
    Cliff's delta & $\hookrightarrow$ & \textbf{0.47} & \textbf{0.66} & \textbf{0.67} & \textbf{0.73} \\ 
\hline

@@

		& none & $\leq 2$ & $[3 - 4]$ & $[5 - 7]$ & $\geq 7$  \\  
		\hline
		\multicolumn{6}{c}{\textbf{Edits}} \\
		\hline
		mean & 3.97 &	2.18 & 1.82 & 1.70 & 1.59 \\ 
		% \cline{3-6} 
		%t.test & $\hookrightarrow$& \multicolumn{4}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		U-test (adjusted) & $\hookrightarrow$ & \multicolumn{4}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		Cliff's delta & $\hookrightarrow$	& \textbf{0.42} & \textbf{0.53} & \textbf{0.57} & \textbf{0.63}    \\

		\hline
		\multicolumn{6}{c}{\textbf{Selections}} \\
		\hline 
		mean & 3.10 &	2.93 & 1.80 & 1.77 & 1.63  \\ 
		% \cline{3-6} 
		%t.test & $\hookrightarrow$& 6.258e-08 & 1.623e-12 & 4.87e-16 & $<$ 2.2e-16 \\
		\cline{3-6} 
		U-test (adjusted)& $\hookrightarrow$ & \multicolumn{4}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		Cliff's delta & $\hookrightarrow$	& \textbf{0.29} & \textbf{0.38} & \textbf{0.39} & \textbf{0.49} \\  
		\hline

		\multicolumn{6}{c}{\textbf{Edit ratio}} \\
		\hline 
		mean & 0.68 & 0.56 & 0.55 & 0.54 & 0.57 \\ 
		\cline{3-6} 
		%    t.test & $\hookrightarrow$& 1.62E-13 & \multicolumn{3}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		U-test (adjusted) &  $\hookrightarrow$& \multicolumn{4}{c}{$<$ 0.0001} \\
		\cline{3-6} 
		Cliff's delta & $\hookrightarrow$ & \textbf{0.19} & \textbf{0.24} & \textbf{0.28} & \textbf{0.26} \\ 
		\hline

@@

\begin{table}[ht!]
\tiny
\renewcommand{\arraystretch}{1.3}
\caption{Effect size of the ratio of edits around the interruption.}
\label{tbl:ratio_all}
\centering
\begin{tabular}{l | p{0.7cm} | p{1.9cm} | p{1.9cm} } 
   & [1-5] & [6-10] & [11-15] \\  
  \hline
  \multicolumn{4}{c}{\textbf{before}} \\
  \hline
  mean & 0.34 &	0.37 &	0.39 \\ 
   \cline{3-4} 
  U-test (adjusted) & $\hookrightarrow$ & \multicolumn{2}{c}{$<$ 2.2e-16}  \\

@@

  \multicolumn{4}{c}{\textbf{after}} \\
  \hline 
  mean & 0.34 &	0.37 &	0.39 \\ 
   \cline{3-4} 
  U-test (adjusted) & $\hookrightarrow$ & 4.40e-16 & $<$ 5.67e-16 \\  
  \cline{3-4} 
  Cliff's delta & $\hookrightarrow$	& \textbf{-0.03} & \textbf{-0.02}  \\  
\end{tabular}
\end{table}

@@

\paragraph{The effects of multitasking}
Certainly, work fragmentation is not a synonym of lower productivity rates. Multitasking can allow the worker to use dead periods of the project where there is a decrease of the required activities and do tasks related to other projects, increasing the efficiency and productivity \cite{ABV12}. At some level, switching between tasks can provide the needed arousal and alertness to improve the productivity, but higher levels of multitasking can be detrimental and cause a negative effect. Moreover, as the number of switches between tasks increases, the effectiveness degrades in a linear fashion \cite{AB12}.

Our results indicate a clear drop of productivity when facing interruptions, but with a high number of interruptions the change in productivity stops decreasing. This could be reflecting what observational studies tell us about multitasking, which is that with higher number of interruptions (increased multitasking) the productivity stops being affected and might increase. %The latter actually occurs in the results but the effect is too subtle. 
We inspected  the effects of work fragmentation in more details by classifying the interruptions into negative or positive ones, and in the results we can see a difference in productivity and the kind of activities performed.

>>>>

From these research questions, we got the following findings and conclusions:
So, in short these are our findings and conclusions:
\begin{itemize}
	\item We found an inverse relationship between the number of interruptions and our productivity metrics.
	\item A high proportion of prolonged interruptions is more detrimental to productivity than short interruptions.
	\item The productivity 30 minutes around an interruption is lower than the average.
	\item There are two well-defined patterns around the interruption, depending if the effect afterwards was positive or negative.
	\item There is a peak of productivity after the recovery time that lasts until approximately 10 minutes before the next interruption.
	\item There is a difference between the activities performed after an interruption, depending if it was positive or negative.
\end{itemize}
%\RR{For each research question, add a sentence explaining why we look into it}

@@

   & none & $\leq 3$ & $[4 - 5]$ & $[6 - 10]$ & $\geq 11$  \\  
  \hline
  \multicolumn{6}{c}{\textbf{Edits}} \\
  \hline
  median & 6.56 & 1 & 0.67 & 0.65 & 0.6	 \\
  \cline{2-6} 
  mean & 6.29 &	2.59 & 1.55 & 1.29 & 0.91  \\ 
  % \cline{3-6} 
  %t.test & $\hookrightarrow$& \multicolumn{4}{c}{$<$ 2.2e-16} \\
   \cline{3-6} 
  U-test (adjusted) & $\hookrightarrow$ & \multicolumn{4}{c}{$<$ 2.2e-16} \\
  \cline{3-6} 
  Cliff's delta & $\hookrightarrow$	& \textbf{0.61} & \textbf{0.77} & \textbf{0.80} & \textbf{0.87}    \\
  \hline

  \multicolumn{6}{c}{\textbf{Selections}} \\
  \hline 
  median & 4.56 & 2.08 & 1.71 & 1.42 & 1.28 \\
  \cline{2-6} 
  mean & 4.73 &	2.93 & 2.26 & 1.79 & 1.54  \\ 
    % \cline{3-6} 
    %t.test & $\hookrightarrow$& 6.258e-08 & 1.623e-12 & 4.87e-16 & $<$ 2.2e-16 \\
     \cline{3-6} 
    U-test (adjusted) & $\hookrightarrow$ & 8.1e-13 & \multicolumn{3}{c}{$<$ 4.4e-16} \\

@@

  \multicolumn{6}{c}{\textbf{Edit ratio}} \\
  \hline 
  median & 0.56 & 0.36 & 0.30 & 0.32 & 0.33 \\
  \cline{2-6} 
  mean & 0.55 & 0.39 & 0.33 & 0.34 & 0.33 \\ 
  \cline{3-6} 
%    t.test & $\hookrightarrow$& 1.62E-13 & \multicolumn{3}{c}{$<$ 2.2e-16} \\
     \cline{3-6} 
    U-test (adjusted) & $\hookrightarrow$ & 4.2e-14& \multicolumn{3}{c}{$<$ 2.9e-16} \\
    \cline{3-6} 
    Cliff's delta & $\hookrightarrow$ & \textbf{0.47} & \textbf{0.66} & \textbf{0.67} & \textbf{0.73} \\ 
\hline

@@

		& none & $\leq 2$ & $[3 - 4]$ & $[5 - 7]$ & $\geq 7$  \\  
		\hline
		\multicolumn{6}{c}{\textbf{Edits}} \\
		\hline
		median & 3.58 & 1.5 & 1.23 & 1.15 & 1.18 \\
		\cline{2-6} 
		mean & 3.97 &	2.18 & 1.82 & 1.70 & 1.59 \\ 
		% \cline{3-6} 
		%t.test & $\hookrightarrow$& \multicolumn{4}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		U-test (adjusted) & $\hookrightarrow$ & \multicolumn{4}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		Cliff's delta & $\hookrightarrow$	& \textbf{0.42} & \textbf{0.53} & \textbf{0.57} & \textbf{0.63}    \\

		\hline
		\multicolumn{6}{c}{\textbf{Selections}} \\
		\hline 
		median & 2.54 & 1.66 & 1.5 & 1.5 & 1.4 \\
		\cline{2-6} 
		mean & 3.10 &	2.93 & 1.80 & 1.77 & 1.63  \\ 
		% \cline{3-6} 
		%t.test & $\hookrightarrow$& 6.258e-08 & 1.623e-12 & 4.87e-16 & $<$ 2.2e-16 \\
		\cline{3-6} 
		U-test (adjusted)& $\hookrightarrow$ & \multicolumn{4}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		Cliff's delta & $\hookrightarrow$	& \textbf{0.29} & \textbf{0.38} & \textbf{0.39} & \textbf{0.49} \\  
		\hline

		\multicolumn{6}{c}{\textbf{Edit ratio}} \\
		\hline 
		median & 0.57 & 0.49 & 0.45 & 0.43 & 0.44 \\
		\cline{2-6} 
		mean & 0.68 & 0.56 & 0.55 & 0.54 & 0.57 \\ 
		\cline{3-6} 
		%    t.test & $\hookrightarrow$& 1.62E-13 & \multicolumn{3}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		U-test (adjusted) &  $\hookrightarrow$& \multicolumn{4}{c}{$<$ 0.0001} \\
		\cline{3-6} 
		Cliff's delta & $\hookrightarrow$ & \textbf{0.19} & \textbf{0.24} & \textbf{0.28} & \textbf{0.26} \\ 
		\hline

@@

\begin{table}[ht!]
\tiny
\renewcommand{\arraystretch}{1.3}
\caption{Effect size of the ratio of edits around the interruption.}
\label{tbl:ratio_all}
\centering
\begin{tabular}{l | p{0.7cm} | p{1.9cm} | p{1.9cm} } 
   & [1-5] & [6-10] & [11-15] \\  
  \hline
  \multicolumn{4}{c}{\textbf{before}} \\
  \hline
  median & 0.35 & 0.37 & 0.37 \\
  \cline{2-4} 
  mean & 0.34 &	0.37 &	0.38 \\ 
   \cline{3-4} 
  U-test (adjusted) & $\hookrightarrow$ & \multicolumn{2}{c}{$<$ 2.2e-16}  \\

@@

  \multicolumn{4}{c}{\textbf{after}} \\
  \hline 
  median & 0.36 & 0.37 & 0.37 \\
  \cline{2-4} 
  mean & 0.34 &	0.37 &	0.39 \\ 
   \cline{3-4} 
  U-test (adjusted) & $\hookrightarrow$ & 4.40e-16 & $<$ 5.67e-16 \\  
  \cline{3-4} 
  Cliff's delta & $\hookrightarrow$	& \textbf{-0.03} & \textbf{-0.02}  \\  
\end{tabular}
\end{table}

@@

\paragraph{The effects of multitasking}
Certainly, work fragmentation is not a synonym of lower productivity rates. Multitasking can allow the worker to use dead periods of the project where there is a decrease of the required activities and do tasks related to other projects, increasing the efficiency and productivity \cite{ABV12}. At some level, switching between tasks can provide the needed arousal and alertness to improve the productivity, but higher levels of multitasking can be detrimental and cause a negative effect. Moreover, as the number of switches between tasks increases, the effectiveness degrades in a linear fashion \cite{AB12}. Also, it was observed that switching between projects might as well degrade the productivity levels of the programmer due to the context switches, so developers that work on many projects per day have lower productivity \cite{VBX16}.

Our results indicate a clear drop of productivity when facing interruptions: when we increase the number of interruptions, the productivity metrics tend to decrease. This agrees with the observational studies cited before.
%Our results indicate a clear drop of productivity when facing interruptions, but with a high number of interruptions the change in productivity stops decreasing. This could be reflecting what observational studies tell us about multitasking, which is that with higher number of interruptions (increased multitasking) the productivity stops being affected and might increase. %The latter actually occurs in the results but the effect is too subtle. 
%We inspected the effects of work fragmentation in more detail by classifying the interruptions into negative or positive ones, and in the results we can see a difference in productivity and the kind of activities performed.
